### 2018-03-09 REBUTTAL.md
Addressing RW1 comments

We want to thank the reviewers for their well-intentioned and valuable
comments. We describe here how we change the manuscript to adress their
comments.

=== RW1 ===

> There's a small typo in the code README, last section on parallelization
> reads 'multiple CPU cores a moder computer provide'.

We corrected the typo in the README file.

> In Figure 2 (of the replication), I appreciate the switch from the
> epsilon+subscript axis labels used in the paper to plain text, which avoids
> having to read the caption. In line with that, I think using full text for
> inhibitory and excitatory (instead of Inh and ext) would be even clearer.

We modified the axis name to have full text inhibitory et excitatory

> Apologies if I missed it, but I don't understand why the replication of
> Figure 3c is skipped. I understand that the crux of the matter is captured by
> a and b, but was there a particular reason why it was omitted (e.g.
> computation time)?  We added the third panel.

We added the third panel

> I would encourage, in the concluding Discussion section, a brief recap of the
> differences found (detail in Figure 2, underestimation in Figure 3) before
> discussing the potential causes.

We modified the discussion accordingly (see L434 in the .tex file)

> I haven't gone through the details of the derivation of the semi-analytic
> solution yet, but I think it would be useful to have a sentence or two
> specifying if/where the derivation presented differs from the original one.
> I don't fully agree with the the statement that the semi-analytical result
> underestimates the numerical one, if anything to me it looks shifted to the
> right (underestimates before the peak, overestimates after).  We corrected
> the text, the difference stem wgqwweee think from the estimation of voltage
> distribution.  I am slightly confused by how you obtain a numerical
> approximation of P(V) as opposed to how the original paper does it.

The numerical curve are in fact shifted upwards due to false positve (see
L430). This is due to false positive spikes.  The difference lays in how we
sample the voltage distribution otherwise we did a similar computation. We
acknowledge that details in the original work were unclear to reproduce the
analytic approach (see L333)

> From what I understand, you build 50 linear networks and 50
> nonlinear networks (with random connectivity )and simulate each network once
> for 250ms (but from how it's written it could sound as though you are running
> multiple simulations of the same two networks - i.e. with the same
> connections), whereas in the original they build 100 random networks (50 and
> 50) and simulate each ten times. Could you clarify what simulations you are
> running to estimate P(V) and how they compare to what the original paper does?
> It may also be good to address this difference in the paper, as you have done
> in most other cases.  We clarified this point in the main text Edit: I think my
> confusion in the point above may also be partly due to the fact that when you
> say 'a network is simulated n times' you mean that also new random connectivity
> is generated in each simulation (whereas for me intuitively simulating a
> network n times means keeping the architecture and drawing new initial
> potentials). This becomes clear since elsewhere you say: 'Each network was
> simulated 10 times with different initial conditions (synaptic connections and
> initial membrane potential values)'. It may be good to be explicit for clarity

We now further clarify this point.

=== RW2 ===

> Major issue: First, while many of the
> decisions made by the authors have been explained, such as the choice of Brian
> and certain parameter adjustments, I am missing an explanation for why the
> authors decided not to use the original analytical solution as well.
> Furthermore, differences between this Figure (3) and the original that are
> described are too short. Instead of only replicating the figure and describing
> a "good match", I would like to see a more quantitative comparison (e.g.,
> position of the maximum, intersections, etc.). The same applies to Fig. 2 (see
> below).  Minor issues: Manuscript: Nice introduction. In line 2, there should
> be a comma after "However". Maybe, a final sentence could summarize the main
> findings of the replication.  Methods are very intuitive and sufficiently
> described. On page 1, I would suggest removing "dynamic" after membrane
> potential. Also, in my personal opinion, often used comments in brackets
> disturb the flow of reading.  In Eq. 1, I would reformat the brackets encasing
> the nonlinear function to be larger.  On page 2, there is a period mark missing
> when describing the spike train function \epsilon.  Why is the equation for the
> nonlinearity function not numbered? I would also suggest moving the commas
> after the cases. After "otherwise" there should be a period mark.  When
> describing the biological observation of nonlinearity only affecting
> synchronous spikes, I would like to see references.  Fantastic description of
> what the neuron model effectively is (point neuron with a single nonlinear
> dendrite) I enjoyed reading the argumentation for a clock-driven model with
> finite precision in terms of biological plausibility.  The description for
> implementing the nonlinear networks could be rewritten to match closer the
> mathematical definitions. Start with describing the use of the clipping
> function to achieve the nonlinearity function, then explain that ve is used as
> a placeholder variable equivalent to x from the equation.  Nice description and
> use of the joblib package.  In describing the grid search, I would write: "For
> the grid search presented in Fig. 2, we varied inhibitory and excitatory
> synaptic weights between 0.16 mV and 0.4 mV. Given that each neuron receives on
> average 150 excitatory and 150 inhibitory inputs, this corresponds to an
> average total weight between [...]" Below on page 4, there is a space missing
> for "50 ms".  The derivation of the semi-analytical approach is well derived
> and described sufficiently.  I would add to the Eq. 6 description, that it is
> the expected value of spiking neurons in the next iteration given a certain
> initial synchronous pulse (g_i^').  Results section describes the similarities
> between original and replication sufficiently, but mainly in qualitative terms.
> In Fig. 1, I would suggest using similar yticks for the rate and the number of
> synchronous spikes, as well as the xticks for the time. This allows a much
> better quantitative evaluation of the reproducibility of the original study.
> In Fig. 2, the description of the resulting color code is very elucidating, as
> this was not very well explained in the original study.  I would like to
> suggest to explain, why the C panel from the original study has not been
> replicated and further the differences could be described in more detail. I
> also see differences for large weights (both exc. and inh.), as well as
> qualitative differences for the linear network (no yellow line/areas). In B,
> the area of unstable activity before stimulation seems to be significantly
> larger.  Why did the authors not try the initial random spikes in transit?
> Fig. 3 I discussed in the major point above.  Code: Well written, a bit
> under-commented, but clearly structured The easy installation of requirements
> and use of CL options make the code very reusable and in general user-friendly
> In lib.py, I am a bit puzzled about the way how the function "run_with_cache"
> is defined. Why can't the run function be decorated with @mem.cache as for the
> other functions?  Could the authors elaborate on the definition of V0 in line
> 81 . It does not seem to be included in Table 1 or the main text.  On first run,
> I got thrown a ton of FutureWarnings concerning the use of certain dtypes in
> Brian. I think this is related to the version/system I was using (although I
> used the requirements.txt to get the exact version numbers). I omitted the
> warnings by using the Warnings library to ignore any FutureWarnings I then
> performed the simulations using the --long option, which finished after 7h38m.